Tables

### 2020-12-18 tables.Rmd
#11 add private HS universe graph

>>>>

# Private HS universe

```{r}
privhs_universe <- readRDS('./data/privhs_universe.RDS')

# Filter for private HS that fits our criteria
# https://github.com/ksalazar3/recruiting-static-visuals/blob/62b08a202ca7c93433a91d7e72e6c0e93ac6eb5b/sql/update_imported_hs_data.sql#L52-L56
val_labels(privhs_universe$school_type)
privhs_universe %>% group_by(year) %>% count()

v_get_val <- Vectorize(function(x) str_to_lower(val_label(privhs_universe$region, x)))

df <- privhs_universe %>%
  filter(school_type %in% c(1, 2, 5), total_12 > 10) %>% 
  mutate(
    region = v_get_val(region),
    pct_blacklatinxnative = pct_black + pct_hispanic + pct_amerindian + pct_nativehawaii,
    pct_blacklatinxnative_cat = case_when(
      pct_blacklatinxnative < 10 ~ 'c1_lt10',
      pct_blacklatinxnative < 25 ~ 'c2_10to25',
      pct_blacklatinxnative < 50 ~ 'c3_25to50',
      pct_blacklatinxnative >= 50 ~ 'c4_50+'
    ),
    rank_cat2 = case_when(
      rank_within_category <= 200 & overall_niche_letter_grade == 'A+' ~ 'c1_top200',
      rank_within_category > 200 & overall_niche_letter_grade == 'A+' ~ 'c2_A+',
      overall_niche_letter_grade == 'A' ~ 'c3_A',
      !is.na(overall_niche_letter_grade) ~ 'c4_ltA'
    ) 
  ) %>% 
  select(total_12, region, religion, pct_blacklatinxnative_cat, rank_cat2)

plot_characteristic <- function(var_name, type, characteristic, label_text) {

  sub_df <- df %>% group_by(get(var_name)) %>% 
    summarize(unweighted = n(), weighted = sum(total_12)) %>% 
    pivot_longer(cols = c('unweighted', 'weighted'),
                 names_to = 'universe',
                 values_to = 'value')
  names(sub_df) <- c('key', 'universe', 'value')
  sub_df <- sub_df %>% filter(!is.na(key))

  sub_df$universe <- factor(sub_df$universe, levels = c('weighted', 'unweighted'))
  sub_df$key <- factor(sub_df$key, levels = characteristic)

  ggplot(sub_df, aes(fill = key, y = value, x = universe)) + 
      geom_bar(stat='identity', position = position_fill(reverse = TRUE), width = 0.7) +
      ggtitle('Title') +
      xlab('') + ylab('') +
      labs(fill = type) +
      scale_fill_manual(labels = label_text, values = c('#F8766D', '#7CAE00', '#00BFC4', '#C77CFF')) +
      scale_y_continuous(labels = scales::percent, expand = c(0, 0, 0.05, 0)) +
      theme(legend.position = 'top',
            legend.title = element_text(size = 6, face = 'bold'),
            legend.text = element_text(size = 6),
            legend.margin = margin(0, 0, 0, 0),
            legend.box.margin = margin(0, 0, -5, -30),
            legend.key.size = unit(0.3, 'cm'),
            panel.background = element_blank(),
            axis.ticks.y = element_blank(),
            text = element_text(size = 8),
            plot.title = element_text(color = 'white')) +
      coord_flip()
}

pdf(file.path('.', 'assets', 'figures', 'plot_pss_universe.pdf'), width = 8, height = 4)
par(mar=c(0, 0, 0, 0) + 0.1, mai=c(0, 0, 0, 0))
par(mfrow=c(2, 2))

a <- plot_characteristic('region', 'Region', c('northeast', 'midwest', 'south', 'west'), c('Northeast', 'Midwest', 'South', 'West'))
b <- plot_characteristic('religion', 'Religion', c('catholic', 'christian', 'other', 'nonsectarian'), c('Catholic', 'Christian', 'Other religion', 'Nonsectarian'))
c <- plot_characteristic('pct_blacklatinxnative_cat', '% Black/Latinx/Native', c('c1_lt10', 'c2_10to25', 'c3_25to50', 'c4_50+'), c('<10%', '10-25%', '25-50%', '50%+'))
d <- plot_characteristic('rank_cat2', 'Ranking', c('c1_top200', 'c2_A+', 'c3_A', 'c4_ltA'), c('Top 200 A+', 'Other A+', 'A', 'Less than A'))

grid.arrange(a, b, c, d, nrow=2, ncol=2)
dev.off()
```
